Transmembrane helices are also an overlooked source of major histocompatibility complex class II epitopes and evolutionary more conserved than expected by chance

### 2021-03-24 bbbq_article.tex
Process feedback Frans

\frans{Show off to Frans}
Cytolytic T cell responses are predicted to be biased towards membrane proteins. 
Because the peptide-binding grooves of most haplotypes 
of histocompatibility complex class I (MHC-I) are relatively hydrophobic, 
peptide fragments derived from transmembrane helices (TMHs) 
are predicted to be presented more often than expected,
based on their abundance in the proteome.
We show that the over-presentation of TMH-derived peptides is likely general, 
as it is predicted for diverse microbiota and viruses 
and for both MHC-I and MHC-II
However, the physiological reason of why membrane proteins might be 
over-presented is unclear
In this study, we show that TMHs are evolutionarily more conserved, 
as relatively less single nucleotide polymorphisms (SNPs) 
are present in TMH-coding chromosomal regions 
compared to regions coding for extracellular and cytosolic protein regions. 
Thus, our findings suggest that T cells might respond 
more to membrane proteins, because these are evolutionary more conserved.
We speculate that TMHs therefor might be less prone to escape mutations 
that enable pathogens to evade T cell responses.

@@

Per protein, we calculated two percentages: 
(1) the percentage of the protein coding for TMHs, 
and (2) the percentage of SNPs occurring in TMHs.
Each percentage pair was plotted in figure \ref{fig:f_snps_found_and_expected}.
We performed a linear regression analyses on the data twice,
one with and one without MAPs,
and added a 95\% confidence interval for both of these regressions.
Would the confidence interval of the regression line 
overlap with the line of equality,
this would be expected if SNPs occur just as likely within TMHs 
and soluble protein domains.
This analysis gives the first hint that SNPs 
occur more often in soluble domains than in TMHs.\frans{waar vind ik dan nu het verschil tussen met en zonder MAPs je zou hopen dat er geen verschil is toch omdat het aandeel TMH in de analyse met MAPS ook naar beneden gaat}
To statistically test whether SNPs occur 
just as likely in soluble domains as
in TMHs, we first calculated the number of SNPs that are expected to be
found in TMHs of TMPs by chance, $E(n_{SNP\_TMH})$,
equals the number of SNPs in TMHs, $n_{SNP} = 21,738$, 
multiplied by the fraction of the TMPs that are TMH, $f_{TMH} = 18.66\%$,
$E(n_{SNP\_TMH}) = n_{SNP} \cdot f_{TMH} \approx 4056$.

The effect size is that, 
when randomly selecting SNPs, 
per 100 SNPs found in soluble protein domains, 
one finds 91 SNPs in TMHs,
as depicted in figure \ref{fig:conservation}.

@@

We conclude that,
when randomly selecting SNPs, 
for 100 SNPs found in soluble regions of multi-spanners,
one finds 89.6 SNPs in TMHs, as shown by \frans{it says 91,1 in fig 4}
figure \ref{fig:conservation_per_spanner}.

@@

[GVDB: 
  Discussion section needs some work, 
  as here interpretation of the results needs to be given. 
  What is the consequence of the findings? Why is it important. 
]

@@

Our understanding of the immune system and epitope prediction is 
important to fight disease. By this paper, we suggest that epitopes
derived from TMHs are needlessly overlooked and may inspire to investigate
new potential targets for combating pathogens.
\frans{ik zie hier verder geen argumentatie voor deze statement}
\subsection{Elution studies}

@@

The over representation of TMH on MHC-II haplotypes such as HLADQA!*0101/DQB*0501 is more pronounced as for MHC-I with a third of the predicted binders from the human and MtB proteome stem from TMH  human reference proteome and MtB reference proteome to be presented on MHC-II. \frans{ik zoek nog naar associaties}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{IC50s per haplotype}
\label{subsec:ic50s_per_haplotype}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

@@

>>>>

Cytolytic T cell responses are predicted to be biased towards membrane proteins. 
Because the peptide-binding grooves of most haplotypes 
of histocompatibility complex class I (MHC-I) are relatively hydrophobic, 
peptide fragments derived from transmembrane helices (TMHs) 
are predicted to be presented more often than expected,
based on their abundance in the proteome.
We show that the over-presentation of TMH-derived peptides is likely general, 
as it is predicted for diverse microbiota and viruses 
and for both MHC-I and MHC-II
However, the physiological reason of why membrane proteins might be 
over-presented is unclear
In this study, we show that TMHs are evolutionarily more conserved, 
as relatively less single nucleotide polymorphisms (SNPs) 
are present in TMH-coding chromosomal regions 
compared to regions coding for extracellular and cytosolic protein regions. 
Thus, our findings suggest that T cells might respond 
more to membrane proteins, because these are evolutionary more conserved.
We speculate that TMHs therefor might be less prone to escape mutations 
that enable pathogens to evade T cell responses.

@@

Per protein, we calculated two percentages: 
(1) the percentage of the protein coding for TMHs, 
and (2) the percentage of SNPs occurring in TMHs.
Each percentage pair was plotted in figure \ref{fig:f_snps_found_and_expected}.
We performed a linear regression analysis on the TMPs
and added a 95\% confidence interval on this regression.
Would the confidence interval of the regression line 
overlap with the line of equality,
this would be expected if SNPs occur just as likely within TMHs 
and soluble protein domains.
This analysis gives the first hint that SNPs 
occur more often in soluble domains than in TMHs.
To statistically test whether SNPs occur 
just as likely in soluble domains as
in TMHs, we first calculated the number of SNPs that are expected to be
found in TMHs of TMPs by chance, $E(n_{SNP\_TMH})$,
equals the number of SNPs in TMHs, $n_{SNP} = 21,738$, 
multiplied by the fraction of the TMPs that are TMH, $f_{TMH} = 18.66\%$,
$E(n_{SNP\_TMH}) = n_{SNP} \cdot f_{TMH} \approx 4056$.

The effect size is that, 
when randomly selecting SNPs,
within TMPs that are 50\% TMH (or correcting for this percentage),
per 100 SNPs found in soluble protein domains, 
one finds 91 SNPs in TMHs,
as depicted in figure \ref{fig:conservation}.

@@

We conclude that,
when randomly selecting SNPs, 
within TMPs that are 50\% TMH (or correcting for this percentage),
for 100 SNPs found in soluble regions of multi-spanners,
one finds 91.1 SNPs in TMHs, as shown by
figure \ref{fig:conservation_per_spanner}.

@@

@@

Our understanding of the immune system and epitope prediction is 
important to fight diseases. By this paper, we suggest that epitopes
derived from TMHs are needlessly overlooked and may inspire to investigate
new potential targets for combating pathogens. We say 'needlessly',
as TMH-derived epitopes are predicted to be 
presented more often than expected by chance, 
combined with are finding that TMHs are evolutionary conserved.

\subsection{Elution studies}

@@

The over-presentation of TMHs for MHC-II haplotypes 
is more pronounced than for MHC-I,
as with a third of the predicted binders 
HIERO
from the human and MtB proteome stem from TMH human reference proteome and MtB reference proteome to be presented on MHC-II. \frans{ik zoek nog naar associaties}

such as \verb;HLA-DQA1*0101/DQB1*0501;

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{IC50 values of binders per haplotype}
\label{subsec:ic50s_per_haplotype}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Presentation of TMH-derived epitopes}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

Tables \ref{tab:tmh_binders_mhc1} and \ref{tab:tmh_binders_mhc2}
show the exact number of binders, binders that overlap with TMHs
and the percentage of binders that overlap with TMHs, as
visualized by figures \ref{fig:1} and \ref{fig:2} respectively.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Relative presentation of TMH-derived epitopes}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

To compare the over-presentation of TMH-derived epitopes between the
different proteomes, we normalized this percentages in such a
way that 1.0 is the percentage of TMH-derived epitopes that would 
be expected by chance. 
Figure \ref{fig:f_tmh_mhc1_normalized} and \ref{fig:f_tmh_mhc2_normalized}
show these normalized values for the MHC-I and MHC-II haplotypes respectively.

@@

To determine the additional over-presentation of TMH-derived epitopes 
in MHC-II (as compared to MHC-I), we used the same normalized percentages
for the human proteome only and compared the averages between MHC-I and MHC-II.
See figure \ref{fig:fig_rel_presentation}.

\begin{figure}[!htbp]
  \includegraphics[width=\textwidth]{bbbq_1_smart_results/fig_rel_presentation.png}
  \caption{
    Normalized proportion of MHC-I and MHC-II epitopes overlapping with TMHs
    for human proteomes.
  }
  \label{fig:fig_rel_presentation}
\end{figure}
